Automatic discovery of cell types and microcircuitry from neural connectomics

### 2014-12-10 docs/paper.tex
arxiv sub
elife sub
lots of updates need to fix

%\documentclass{nature}
\documentclass{article}
\usepackage[american]{babel}
\usepackage[backend=biber,style=nature]{biblatex}
\usepackage{todonotes}
\usepackage{grffile}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{fullpage}
%\usepackage{algorithm}
%\usepackage{algorithmic}
\usepackage{amsmath}
%\usepackage{subcaption}
%\usepackage{multicol}
\usepackage{csquotes}
\usepackage{pdfpages}
\usepackage{placeins}

@@

\section*{Results}

@@

We apply our algorithm to datasets from mouse
retina, C. elegans and a historical microprocessor. Anatomists
classify cells based on many features, giving us a meaningful baseline
to compare against.

@@

To validate our model, we performed a series of simulations to test if
the model can accurately recover the true underlying network structure
and cell type identity.  We thus simulate data for which we know the
correct structure and comparing the estimated structure based on the
algorithm (see methods) with the one we used for simulation. We find
that the model does a good job of recovering the correct number of
cell types, (fig~\ref{fig:synthetic}a), the the cell identities
(fig~\ref{fig:synthetic}b), and the spatial extent of each type
(fig~\ref{fig:synthetic}c).  For comparison, existing infinite stochastic block model assumes cell type alone matters, and thus finds small neighborhoods of connected nodes (instead of global connectivity patterns). The model converges
relatively quickly to an estimate of the most probable values for the
cell types, which is enabled by using a combination of simulated
annealing and parallelized Markov chain Monte Carlo (see methods for
details). Thus we can apply our model to simulated datasets with structure
and scale similar to that of our biological datasets and recover the known 
correct structure.

@@

Inference is performed in three steps via composable transition 
kernels -- one for structural, one for per-type parameters, and
one kernel for global parameters and hyperparameters. Details
of data preprocessing, inference parameters, and runtime can
be found in the Methods section.

@@

\newpage
\section*{Methods }

>>>>

\documentclass{article}
\usepackage[american]{babel}
\usepackage[backend=biber,style=nature]{biblatex}
\usepackage{fixme}
\usepackage{grffile}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{fullpage}
%\usepackage{algorithm}
%\usepackage{algorithmic}
\usepackage{amsmath}
%\usepackage{subcaption}
%\usepackage{multicol}
\usepackage{csquotes}
\usepackage{pdfpages}
\usepackage{placeins}

@@

\section*{Methods}

@@

As an illustrative example, consider a network with only three cell
types, labeled A, B, and C. Assume the network has three types of
cells, they are uniformly distributed in space, and that the
probability of connection between any two cells, $c_i$ and $c_j$,
depends only on their type and their distance, according to a logistic
(sigmoidal) function. Let A connect only to nearby B and C cells, but B
connect to any C regardless of distance.

Inference is performed in three steps via composable transition 
kernels -- one for structural, one for per-type parameters, and
one kernel for global parameters and hyperparameters. Details
of data preprocessing, inference parameters, and runtime can
be found in the Supplementary Methods section.

@@

\fxwarn{Do we talk about cross-validation early on? What about posterior uncertainty?}

\subsection{Metrics}
We focus on two questions we can ask from the learned model: which
cells are of the same type? and given any two cells, what is the
probability of their connection?

To assess the accuracy of a clustering compared to that determined by
neuroanatomists, we employ the adjustded Rand index (ARI). ARI
measures the similarity of two clusterings. While there are no perfect
clustering comparison metrics (see supplemental for a discussion of
ARI's shortcomings) all suffer from the need to balance
``overclustering'' (calling a single type two different types) with
``merging'' (calling two distinct types a single type).

To test whether our model accurately captures the true structure of
the data, we learn the model on the data with a subset of the links
missing and then compute our predictive accuracy. We perform 10-way
cross validation on a given dataset, learn the resulting model, and
use that model to predict the missing synapses. Each potential
link between cells is assigned a probability, and we compute
the area under the curve (AUC) for the resulting ROC curve. 
\fxwarn{Add additional ARI} 
\section*{Results}

\subsection{Ground Truth Validation}
To validate our model, we performed a series of simulations to test if
the model can accurately recover the true underlying network structure
and cell type identity.  We thus simulate data for which we know the
correct structure and comparing the estimated structure based on the
algorithm (see methods) with the one we used for simulation. We find
that the model does a good job of recovering the correct number of
cell types, (fig~\ref{fig:synthetic}a), the the cell identities
(fig~\ref{fig:synthetic}b), and the spatial extent of each type
(fig~\ref{fig:synthetic}c).  For comparison, existing infinite
stochastic block model assumes cell type alone matters, and thus finds
small neighborhoods of connected nodes (instead of global connectivity
patterns). The model converges relatively quickly to an estimate of
the most probable values for the cell types, which is enabled by using
a combination of simulated annealing and parallelized Markov chain
Monte Carlo (see methods for details). Thus we can apply our model to
simulated datasets with structure and scale similar to that of our
biological datasets and recover the known correct structure.

\subsection{Model Mismatch}
Our type-and-distance assumptions for our connectivty model are
simplifying, and reflect a general consensus about the organization of
some (but not all) types of neural tissue. To show where our model
performs poorly, we attempt connectivity inference on five sets of
synthetic data. Our goal is to show that our model does not find
``structure'' when there is none. \fxwarning{Fix this phrasing -- we want to say 'if your model thinks there are types, there probably are!}.

\begin{itemize}
\item Data with type-specific connectivity and distance dependence
\item Data with type-specific connectivity and no distance dependence
\item Data generated according to a latent-position clustering model,
  where each cell belongs to a type which is part of a cluster in a
  low-dimensional space \fxwarning{Cite this paper}
\item cell type is combinatorial -- there are three possible
  properties a cell can have, and each subset of properties influences
  connectivity to neighboring cells. 
\item cell type is a continuum -- each cell's type continuously varies
  from zero to one, and the probability of a connection between cells
  depends on this type.
\end{itemize}

We generate ten synthetic datasets based on each of these assumptions,
and perform 10-way cross validation to predict link accuracy as well
as ARI to the known (``true'') type, type spatial extent, and type
number.

\fxwarn{Add results}

\subsection{Sensitivity to Edge Effects}

Initial connectomic efforts reconstruct only a small section of tissue
-- as a result, many connections to cells outside that tissue volume
will be lost. We performed simulations to check if 1. edge effects
could destroy spatial structure and if edge effects could introduce
artifical, spurious spatial structure.

\fxwarn{Add ``intrudoce spatial structure where there is none}

@@

Inference is performed in three steps via composable transition 
kernels -- one for structural, one for per-type parameters, and
one kernel for global parameters and hyperparameters. Details
of data preprocessing, inference parameters, and runtime can
be found in the Supplementary Methods section.

@@

\newpage
\section*{Supplementary Methods }
